---
title: "Atlantic Storm Visualisation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

This document contains my solutions to the exercises as presented in https://www.r-bloggers.com/specialize-in-geo-spatial-visualizations-with-leaflet-part-2-exercises/.

First we need to load the **leaflet** package and the dataset **atlStorms2005**.

```{r load}
library(tidyverse)
library(leaflet)
library(htmltools)

data = leaflet::atlStorms2005
```

## Exercise 1

Set the default view to longitude -47.4, latitude 39.75 and zoom level 3.
In addition, add a button that will scale back to the default view upon clicking it.
Hint: use the `icon()` function from the shiny package to easily render an icon of your choice for the button.

### Solution

```{r exercise1}
map = data %>%
  leaflet() %>%
  setView(lng = -47.4,
          lat = 39.75,
          zoom = 3) %>%
  addTiles()

map
```

## Exercise 2

Add the lines that represent the storms traces to the map.

### Solution

```{r exercise2}
map %>%
  addPolylines()
```

## Exercise 3

Color each line according to the storm max wind.

### Solution

```{r exercise3}
# Create colour palette based on MaxWind variable
pal = colorNumeric(palette="Oranges", domain=data$MaxWind)

map %>%
  addPolylines(color=~pal(MaxWind))
```

## Exercise 4

Add a legend to the colors you just added.

### Solution

```{r exercise4}
map %>%
  addPolylines(color=~pal(MaxWind)) %>%
  addLegend("topright", pal = pal, values = ~MaxWind,
    title = "Maximum Wind Speed",
    labFormat = labelFormat(suffix = " mph"),
    opacity = 1)
```

## Exercise 5

Upon hovering over a line, change its weight to 10.

### Solution

```{r exercise5}
map %>%
  addPolylines(color=~pal(MaxWind),
               highlightOptions=highlightOptions(weight=10)) %>%
  addLegend("topright", pal=pal, values=~MaxWind, title="Maximum Wind Speed",
    labFormat=labelFormat(suffix=" mph"), opacity=1)
```

## Exercise 6

Upon hovering over a line, show a label with the storm name.

### Solution

```{r exercise6}
map %>%
  addPolylines(color=~pal(MaxWind),
               highlightOptions=highlightOptions(weight=10),
               label=~htmlEscape(Name)) %>%
  addLegend("topright", pal=pal, values=~MaxWind, title="Maximum Wind Speed",
    labFormat=labelFormat(suffix=" mph"), opacity=1)
```

## Exercise 7

Upon clicking a line, show a pop-up with the storm minimum pressure.

### Solution

```{r exercise7}
map %>%
  addPolylines(color=~pal(MaxWind),
               highlightOptions=highlightOptions(weight=10),
               label=~htmlEscape(Name),
               popup=~htmlEscape(MinPress)) %>%
  addLegend("topright", pal=pal, values=~MaxWind, title="Maximum Wind Speed",
    labFormat=labelFormat(suffix=" mph"), opacity=1)
```

## Exercise 8

Save your leaflet map as an HTML file.
Hint: use `htmlwidgets::save()`.

### Solution

```{r exercise8}
library(htmlwidgets)

# Below we find the final map
final_map = map %>%
  addPolylines(color=~pal(MaxWind),
               highlightOptions=highlightOptions(weight=10),
               label=~htmlEscape(Name),
               popup=~htmlEscape(MinPress)) %>%
  addLegend("topright", pal=pal, values=~MaxWind, title="Maximum Wind Speed",
    labFormat=labelFormat(suffix=" mph"), opacity=1)

htmlwidgets::saveWidget(final_map, "map.html")
```
